{\bf {\Large Parameter sensitivity and identifiability for a biogeochemical model of hypoxia in the northern {G}ulf of {M}exico}}

### 2016-08-22 ident_manu.tex
parcats2 updated for new fishtank
explanation of parameter removal

\begin{singlespace}
\title{{\bf {\Large Parameter sensitivity and identifiability for a biogeochemical model of hypoxia in the northern {G}ulf of {M}exico}}}
\author{
  {\bf {\normalsize Marcus W. Beck$^1$, John C. Lehrter$^1$, Lisa L. Lowe$^2$}}
  \\\\{\textit {\normalsize $^1$USEPA National Health and Environmental Effects Research Laboratory}}
  \\{\textit {\normalsize Gulf Ecology Division, 1 Sabine Island Drive, Gulf Breeze, FL 32561}}
	\\{\textit {\normalsize Phone: 850-934-2480, Fax: 850-934-2401}}
	\\{\textit {\normalsize Emails: \href{mailto:beck.marcus@epa.gov}{beck.marcus@epa.gov}, \href{mailto:lehrter.john@epa.gov}{lehrter.john@epa.gov}}}
	\\\\{\textit {\normalsize $^2$Lockheed Martin IS \& GS - Civil, supporting the USEPA}}
	\\{\textit {\normalsize Resarch Triangle Park, NC 27709}}
	\\{\textit {\normalsize Phone: 919-541-3985,}}
	\\{\textit {\normalsize Email: \href{mailto:lowe.lisa@epa.gov}{lowe.lisa@epa.gov}}}
  \vspace{1in} 
  \\ Version Date:   Wed Aug 3 11:12:36 2016 -0500
	}
\date{}
\maketitle
\end{singlespace}
\clearpage

@@

This study evaluates a recently developed hydrodynamic and ecological model that describes horizontal and vertical transport and mixing of state variables relevant for hypoxia in the northern \ac{gom}.  The \ac{cgem} includes elements from the Navy Coastal Ocean Model \citep{Martin00} that describe hydrodynamics on the \ac{lcs} and a biogeochemical model with multiple plankton groups, water-column metabolism, and sediment diagenesis \citep{Eldridge10}.  The hydrodynamic component of \ac{cgem} provides a spatially-explicit description of hypoxia using an orthogonal grid with an approximate horizontal resolution of 1.9 km$^2$ and twenty equally-spaced vertical sigma layers on the shelf (depth $\leq$ 100 m, with additional hybrid layers at deeper depths).  The biogeochemical component includes equations for 36 state variables including six phytoplankton groups (with nitrogen and phosophorus quotas for each), two zooplankton groups, nitrate, ammonium, phosphate, dissolved inorganic carbon, oxygen, silica, and multiple variables for dissolved and particulate organic matter from different sources.  Atmospheric and hydrological boundary conditions described in \citet{Hodur97} and \citet{Lehrter13} are also included in \ac{cgem}.

The core unit of \ac{cgem} is FishTank, a \ac{zerod} model that implements the biogeochemical equations in \citet{Eldridge10} and does not include any form of physical transport (i.e., advection, mixing, or surface flux).  Although FishTank was developed for specific application in \ac{cgem}, it can easily be applied to other hydrodynamic grids. Accordingly, the sensitivity and identifiability analysis described below are informative for both the \ac{lcs} gridded model as well as potential applications to different systems.  The FishTank model provides estimates for the 36 state variables described above using a \ac{zerod} parcel that is uniformly mixed as a closed system.  A set of initial conditions is provided as input to the model that was based on observations of relevant variables obtained from research cruises in April, June, and September 2006 \citet[Table 1 in][]{Murrell14}.

Results from FishTank are based on time-dependent differential equations that describe energy flow between phytoplankton and zooplankton groups as affected by nutrient uptake rates, organic matter inputs and losses, inherent optical properties, sediment diagenesis, and temperature (\citealt{Penta08,Eldridge10}, see appendix in \citetalias{LehrterIR}).  A total of 108 equations are estimated at each time step to return a value for each of the 36 state variables described by the model.  In addition to the initial conditions, 250 parameter values for each of the equations is also supplied at model execution.  These parameters define relationships among fixed effects in the equations and represent ecological properties described by the model that influence hypoxia formation.  Values for each of the parameters were based on estimates from the literature, field or laboratory-based measurements, or expert knowledge in absence of the former.  As such, a sensitivity analysis of parameter values is warranted given that, for example, literature or field-based estimates may not apply under all scenarios or expert knowledge is not completely certain \citep{Refsgaard07}.  The sensitivity of \ac{do} to perturbations of all parameters for the 108 equations was estimated from January 1\textsuperscript{st} to December 31\textsuperscript{st}, 2006 by running FishTank at a timestep of five minutes.  For simplicity, the parameters were grouped into one of six categories based on their respective equations: optics ($n = $ 11 parameters), organic matter (29), phytoplankton (156), temperature (32), and zooplankton (22).  A full description is available as an appendix in \citepalias{LehrterIR}.

@@

Local sensitivity analyses showed that \ac{do} was sensitive to perturbations in 140 of 250 parameters (56\% of total) in FishTank (see \cref{fig:sensalltile} for other state variables). Within each parameter category, \ac{do} was sensitive to four parameters for optics (36\% of all optic parameters, \cref{tab:optsens}), seven for organic matter (24\%, \cref{tab:omsens}), 103 for phytoplankton (66\%, \cref{tab:phytosens}), seven for temperature (22\%, \cref{tab:tempsens}), and 19 for zooplankton (86\%, \cref{tab:zoopsens}). Although \ac{do} had the greatest sensitivity to parameters in the zooplankton category (as percentage of total), the relative effects varied. Among all parameters, sensitivity values ranged from $L1 = $ $3.18\times 10^{-6}$ for \textit{Kcdom} (optics) to $328.35$ for \textit{Qc$_{p1}$} (phytoplankton), whereas average sensitivity among all parameters was $L1 = $ $4.52$.  Within categories, sensitivity ranged from $3.18\times 10^{-6}$ (\textit{Kcdom}) to $1.64$ (\textit{astar490}) for optics, $0.01$ (\textit{KGcdom}) to $2.11$ (\textit{k11}) for organic matter, $1\times 10^{-5}$ (\textit{Ksi$_{p1}$}) to $328.35$ (\textit{Qc$_{p1}$}) for phytopankton, $0.12$ (\textit{Tref(nospA+nospZ)$_{p1}$}) to $2.9$ (\textit{Tref(nospA+nospZ)$_{p4}$}) for temperature, and $0$ (\textit{ZQp$_{z1}$}) to $0.82$ (\textit{ZKa$_{z1}$}) for zooplankton (\cref{fig:sensplo}, bottom).  Average sensitivity values in each category were $L1 = $ $0.42$ for optics, $1.24$ for organic matter, $5.96$ for phytoplankton, $0.55$ for temperature, and $0.27$ for zooplankton.  Within the six phytoplankton groups, \ac{do} was sensitive to the same parameters within each group although the sensitivity magnitudes varied. Average sensitivity across parameters in each phytoplankton groups showed that \ac{do} was most sensitive to the first and third phytoplankton groups (average $L1 = $ $19.71$, $14.56$), whereas sensitivity to parameters in the remaining groups was much lower (all with average $L1 < 1$). Sensitivity of \ac{do} did not vary considerably between parameters in the two zooplankton groups (average $L1 = $ $0.31$, $0.23$ for groups one and two).

Response of \ac{do} to parameter perturbations was not uniform across the time series.  \Cref{fig:sensplo} shows variation for the top parameters within each category.  Because FishTank does not include a spatial component, the estimated \ac{do} trend describes a closed, heterotrophic system where respiration processes eventually remove all \ac{do} from the model space.  The initial decrease in the time series reflects change from the initial conditions outside of the growing seasons (i.e., January, February), the spring/summer increase represents production associated with expected seasonal growth, and the remaining time series from August to the end of year shows complete removal of \ac{do} as respiration processes dominate metabolic activity. Although this time series is not a realistic depiction of an actual system, the biogeochemical model behaves as expected in the absence of the hydrodynamic model.  Accordingly, the interpretation of sensitivity results from the simple model has relevance in an ecological context.  As expected, parameter perturbations had the largest effect during the summer months, although the effects varied.  An inrease in 50\% from the parameter default values generally caused a reduction in \ac{do} during the summer, with the exception of the zooplankton parameter, \textit{ZKa$_{z1}$}, which caused an increase in \ac{do}.  The phytoplankton parameter, \textit{Qc$_{p1}$}, had the largest effect such that the \ac{do} time series was similar to the default output in April/May, whereas a dramatic decrease was observed in the remaining months.  The effects of perturbations early in the time series (January, February) showed similar patterns such that a reduction in \ac{do} was most common, particularly for the optics (\textit{astar490}) and phytoplankton (\textit{Qc$_{p1}$}) parameters.

@@

The identifiability analyses suggested that most parameter subsets exceeded the thresholds of $\gamma = 10, 15$, providing further justification for using selection heuristics for parameter optimization.  Parameter identifiability (as $\gamma$) increased at different rates depending on the parameter category or the number of top parameters that were selected (\cref{fig:identbox}).  By category, identifiability was lowest for parameter subsets in the phytoplankton ($7$\% less than $\gamma = 15$, $5.9$\% less than $\gamma = 10$)  and zooplankton categories ($28.9$\%, less than $\gamma = 15$, $24.2$\%, less than $\gamma = 10$), whereas a majority of combinations for temperature were identifiable ($83.3$\% less than $\gamma = 15$, $73.3$\%, less than $\gamma = 10$).  All subset combinations for optics and organic matter parameters had $\gamma < 10$.  All parameter subsets for choosing the top, top two, and top three parameters in each category were identifiable (\cref{fig:identbox}), whereas a majority were identifiable for choosing the top four ($77$\% less than $\gamma = 15$, $70.8$\% less than $\gamma = 10$) and top five ($77$\% less than $\gamma = 15$, $70.8$\%, less than $\gamma = 10$) parameters.

@@

Each of the three selection heuristics (blocked by parameter category, independent of category, all categories equally) for \ac{do} differed in the number of selected parameters and distribution of parameters within each category (if applicable, \cref{fig:heurist,tab:identall1}).  For the first selection heuristic, all sensitive parameters from the optics ($n = $ $4$, $\gamma =$  $3.7$) and organic matter ($n = $ $7$, $\gamma =$  $1.7$) categories were selected, whereas 23 were selected for phytoplankton ($\gamma = $ $13.7$), five for temperature ($\gamma = $ $11.4$), and five for zooplankton ($\gamma = $ $3.1$).  For the second selection heuristic, 20 parameters were selected ($\gamma = $ $4.6$, with a majority from the phytoplankton category, \cref{tab:identall1}).  For the third heuristic, 18 parameters were selected ($\gamma = $ $9.2$) with equal representation between categories. For the second and third selection heuristics, an individual parameter caused a disproportianate increase in $\gamma$ that forced the selection to stop.  Selection independent of category showed that including \textit{KNO3} caused in increase in $\gamma$ from $4.6$ to $15.6$ and equal selection within cateogies showed that including \textit{Tref(nospA+nospZ)$_{p2}$} caused in increase in $\gamma$ from $9.2$ to $66.2$.  Finally, parameter selection for other state variables (\ac{chla}, \ac{par}, nitrate, ammonium, particulate organic matter, dissolved organic matter, phosphorus) also showed that a limited number of parameters were identifiable (for brevity, only results using the second set of selection heuristics are shown, \cref{fig:heurist_stts,tab:identall2}).  Less than ten parameters were selected for each of the remaining state variables with most selected from the phytoplankton category.  Interestingly, no parameters from the temperature category were selected.

@@

% optics sensitivity
%latex.default(totab, file = "", rowlabel = "Description", caption = cap.val,     caption.loc = "top", rowname = Description, size = tabsize,     label = paste0("tab:", tablab))%
\begin{table}[!tbp]
{\normalsize
\caption{Sensitivities of \ac{do} to perturbation of optics parameters.  Sensitivities are based on a 50\% increase from the default parameter value, where $L1$ and $L2$ summarize differences in model output from the default (see \cref{l1,l2}).  Parameters that did not affect \ac{do} are not shown.\label{tab:optsens}} 
\begin{center}
\begin{tabular}{lllll}
\hline\hline
\multicolumn{1}{l}{Description}&\multicolumn{1}{c}{Parameter}&\multicolumn{1}{c}{Value}&\multicolumn{1}{c}{L1}&\multicolumn{1}{c}{L2}\tabularnewline
\hline
Chla specific absorption at 490 nm&\textit{astar490}&$0.04$&$1.64$&$8.64$\tabularnewline
OMA specific absorption at 490 nm&\textit{astarOMA}&$0.1$&$0.02$&$0.1$\tabularnewline
OMZ specific absorption at 490 nm&\textit{astarOMZ}&$0.1$&$0.01$&$0.01$\tabularnewline
AOP, light attenuation due to CDOM&\textit{Kcdom}&$0$&$3.18\times 10^{-6}$&$1.09\times 10^{-5}$\tabularnewline
\hline
\end{tabular}\end{center}}

@@

% organic matter sensitivity
%latex.default(totab, file = "", rowlabel = "Description", caption = cap.val,     caption.loc = "top", rowname = Description, size = tabsize,     label = paste0("tab:", tablab))%
\begin{table}[!tbp]
{\normalsize
\caption{Sensitivities of \ac{do} to perturbation of organic matter parameters.  Sensitivities are based on a 50\% increase from the default parameter value, where $L1$ and $L2$ summarize differences in model output from the default (see \cref{l1,l2}).  Parameters that did not affect \ac{do} are not shown.\label{tab:omsens}} 
\begin{center}
\begin{tabular}{lllll}
\hline\hline
\multicolumn{1}{l}{Description}&\multicolumn{1}{c}{Parameter}&\multicolumn{1}{c}{Value}&\multicolumn{1}{c}{L1}&\multicolumn{1}{c}{L2}\tabularnewline
\hline
rate constant for nitrification&\textit{k11}&$5$&$2.11$&$16.1$\tabularnewline
O2 concentration that inhibits denitrification&\textit{KstarO2}&$10$&$1.91$&$11.76$\tabularnewline
turnover rate for OM1A and OM1G&\textit{KG1}&$50$&$1.49$&$6.83$\tabularnewline
turnover rate for OM2A and OM2G&\textit{KG2}&$50$&$1.37$&$3.46$\tabularnewline
half-saturation concentration for NO3 used in denitrification&\textit{KNO3}&$10$&$1.24$&$7.98$\tabularnewline
half-saturation concentration for O2 utilization&\textit{KO2}&$10$&$0.56$&$1.92$\tabularnewline
decay rate of CDOM, 1/day&\textit{KGcdom}&$0.01$&$0.01$&$0.03$\tabularnewline
\hline
\end{tabular}\end{center}}

@@

% phytoplankton sensitivity
%latex.default(totab, file = "", rowlabel = "Description, Parameter",     caption = cap.val, caption.loc = "top", rgroup = unique(Description),     n.rgroup = as.numeric(table(Description)), rgroupTexCmd = NULL,     rowname = Parameter, size = tabsize, label = paste0("tab:",         tablab))%
\begin{table}[!tbp]
{\normalsize
\caption{Sensitivities of \ac{do} to perturbation of phytoplankton parameters.  Sensitivities are based on a 50\% increase from the default parameter value, where $L1$ and $L2$ summarize differences in model output from the default (see \cref{l1,l2}).  Parameters that did not affect \ac{do} are not shown.  Subscripts show the phytoplankton or zooplankton group that applies for the parameter. Parameters less than the 75\textsuperscript{th} percentile (0.59) for L1 were removed for brevity.\label{tab:phytosens}} 
\begin{center}
\begin{tabular}{llll}
\hline\hline
\multicolumn{1}{l}{Description, Parameter}&\multicolumn{1}{c}{Value}&\multicolumn{1}{c}{L1}&\multicolumn{1}{c}{L2}\tabularnewline
\hline
{coefficient for non-limiting nutrient}&&&\tabularnewline
~~\textit{aN$_{p1}$}&$1$&$2.7$&$47.26$\tabularnewline
~~\textit{aN$_{p3}$}&$1$&$0.66$&$3.16$\tabularnewline
\hline
{half-saturation constant for n}&&&\tabularnewline
~~\textit{Kn$_{p3}$}&$5.93$&$1.15$&$9.22$\tabularnewline
~~\textit{Kn$_{p4}$}&$1.13$&$0.62$&$2.5$\tabularnewline
\hline
{initial slope of the photosynthesis-irradiance relationship}&&&\tabularnewline
~~\textit{alpha$_{p4}$}&$3.96\times 10^{-16}$&$1.61$&$7.46$\tabularnewline
~~\textit{alpha$_{p3}$}&$6.19\times 10^{-17}$&$1.06$&$3.74$\tabularnewline
~~\textit{alpha$_{p5}$}&$3.87\times 10^{-16}$&$0.81$&$5.54$\tabularnewline
\hline
{minimum n cell-quota}&&&\tabularnewline
~~\textit{QminN$_{p3}$}&$1.27\times 10^{-8}$&$1.53$&$6.4$\tabularnewline
~~\textit{QminN$_{p4}$}&$1.53\times 10^{-10}$&$0.91$&$2.76$\tabularnewline
~~\textit{QminN$_{p1}$}&$6.08\times 10^{-9}$&$0.62$&$4.55$\tabularnewline
\hline
{mortality coefficient}&&&\tabularnewline
~~\textit{mA$_{p3}$}&$0.03$&$2.98$&$14.34$\tabularnewline
~~\textit{mA$_{p4}$}&$0.11$&$1.35$&$7.04$\tabularnewline
\hline
{n-uptake rate measured at umax}&&&\tabularnewline
~~\textit{vmaxN$_{p5}$}&$1.4\times 10^{-9}$&$7.57$&$43.9$\tabularnewline
~~\textit{vmaxN$_{p4}$}&$1.33\times 10^{-9}$&$2.88$&$16.56$\tabularnewline
~~\textit{vmaxN$_{p3}$}&$8.11\times 10^{-8}$&$1.99$&$15.23$\tabularnewline
~~\textit{vmaxN$_{p1}$}&$4.1\times 10^{-8}$&$0.84$&$2.92$\tabularnewline
\hline
{p-uptake rate measured at umax}&&&\tabularnewline
~~\textit{vmaxP$_{p3}$}&$6.15\times 10^{-8}$&$2.23$&$9.89$\tabularnewline
~~\textit{vmaxP$_{p1}$}&$2.68\times 10^{-8}$&$0.74$&$11.41$\tabularnewline
\hline
{phytoplankton basal respiration coefficient}&&&\tabularnewline
~~\textit{respb$_{p4}$}&$0.02$&$3.1$&$26.48$\tabularnewline
~~\textit{respb$_{p3}$}&$0.02$&$2.15$&$6.35$\tabularnewline
~~\textit{respb$_{p5}$}&$0.02$&$0.81$&$8.61$\tabularnewline
\hline
{phytoplankton carbon/cell}&&&\tabularnewline
~~\textit{Qc$_{p1}$}&$1.35\times 10^{-6}$&$328.35$&$2181.05$\tabularnewline
~~\textit{Qc$_{p3}$}&$2.65\times 10^{-6}$&$232.8$&$1574.72$\tabularnewline
~~\textit{Qc$_{p2}$}&$1.68\times 10^{-7}$&$4.08$&$61.25$\tabularnewline
~~\textit{Qc$_{p4}$}&$4.54\times 10^{-8}$&$1.03$&$3.75$\tabularnewline
\hline
{phytoplankton growth respiration coefficient}&&&\tabularnewline
~~\textit{respg$_{p4}$}&$0.1$&$0.98$&$7.37$\tabularnewline
\hline
\end{tabular}\end{center}}

>>>>

\begin{singlespace}
\title{{\bf {\Large Parameter sensitivity and identifiability for a biogeochemical model of hypoxia in the northern {G}ulf of {M}exico}}}
\author{
  {\bf {\normalsize Marcus W. Beck$^1$, John C. Lehrter$^1$, Lisa L. Lowe$^2$}}
  \\\\{\textit {\normalsize $^1$USEPA National Health and Environmental Effects Research Laboratory}}
  \\{\textit {\normalsize Gulf Ecology Division, 1 Sabine Island Drive, Gulf Breeze, FL 32561}}
	\\{\textit {\normalsize Phone: 850-934-2480, Fax: 850-934-2401}}
	\\{\textit {\normalsize Emails: \href{mailto:beck.marcus@epa.gov}{beck.marcus@epa.gov}, \href{mailto:lehrter.john@epa.gov}{lehrter.john@epa.gov}}}
	\\\\{\textit {\normalsize $^2$Lockheed Martin IS \& GS - Civil, supporting the USEPA}}
	\\{\textit {\normalsize Resarch Triangle Park, NC 27709}}
	\\{\textit {\normalsize Phone: 919-541-3985,}}
	\\{\textit {\normalsize Email: \href{mailto:lowe.lisa@epa.gov}{lowe.lisa@epa.gov}}}
  \vspace{1in} 
  \\ Version Date:   Sat Aug 13 13:27:22 2016 -0500
	}
\date{}
\maketitle
\end{singlespace}
\clearpage

@@

This study evaluated a recently developed hydrodynamic and ecological model that describes horizontal and vertical transport and mixing of state variables relevant for hypoxia in the northern \ac{gom}.  The \ac{cgem} includes elements from the Navy Coastal Ocean Model \citep{Martin00} that describes hydrodynamics on the \ac{lcs} and a biogeochemical model with multiple plankton groups, water-column metabolism, and sediment diagenesis \citep{Eldridge10}.  The hydrodynamic component of \ac{cgem} provides a spatially-explicit description of hypoxia using an orthogonal grid with an approximate horizontal resolution of 1.9 km$^2$ and twenty equally-spaced vertical sigma layers on the shelf (depth $\leq$ 100 m, with additional hybrid layers at deeper depths).  The biogeochemical component includes equations for 36 state variables including six phytoplankton groups (with nitrogen and phosophorus quotas for each), two zooplankton groups, nitrate, ammonium, phosphate, dissolved inorganic carbon, oxygen, silica, and multiple variables for dissolved and particulate organic matter from different sources.  Atmospheric and hydrological boundary conditions described in \citet{Hodur97} and \citet{Lehrter13} are also included in \ac{cgem}.

The core unit of \ac{cgem} is FishTank, a \ac{zerod} model that implements the biogeochemical equations in \citet{Eldridge10} and does not include any form of physical transport (i.e., advection, mixing, or surface flux).  Although FishTank was developed for specific application in \ac{cgem}, it can easily be applied to other hydrodynamic grids. Accordingly, the sensitivity and identifiability analysis described below are informative for both the \ac{lcs} gridded model as well as potential applications to different systems.  The FishTank model provides estimates for the 36 state variables described above using a \ac{zerod} parcel that is uniformly mixed as a closed system.  A set of initial conditions is provided as input to the model that was based on observations of relevant variables obtained from research cruises in April, June, and September 2006 \citep[Table 1 in][]{Murrell14}.

Results from FishTank are based on time-dependent differential equations that describe energy flow between phytoplankton and zooplankton groups as affected by nutrient uptake rates, organic matter inputs and losses, inherent optical properties, sediment diagenesis, and temperature (\citealt{Penta08,Eldridge10}, see appendix in \citetalias{LehrterIR}).  A total of 108 equations are estimated at each time step to return a value for each of the 36 state variables described by the model.  In addition to the initial conditions, 251 parameter values for each of the equations are also supplied at model execution.  These parameters define relationships among fixed effects in the equations and represent ecological properties described by the model that influence hypoxia formation.  Values for each of the parameters were based on estimates from the literature, field or laboratory-based measurements, or expert knowledge in absence of the former.  As such, a sensitivity analysis of parameter values is warranted given that, for example, literature or field-based estimates may not apply under all scenarios or expert knowledge is not completely certain \citep{Refsgaard07}.  The sensitivity of \ac{do} to perturbations of all relevant parameters for the 108 equations was estimated from January 1\textsuperscript{st} to December 31\textsuperscript{st}, 2006 by running FishTank at a timestep of five minutes.  Initial screening removed 73 irrelevant parameters that were not relevant such that 178 were retained for the sensitivity analysis.  Irrelevant parameters were removed for several reasons, e.g., a parameter was only relevant for the hydrodynamic model, was considered a physical constant, or had no effect given initial conditions.  For simplicity, the remaining parameters were grouped into one of six categories based on their respective equations: optics ($n = $ 4 parameters), organic matter (12), phytoplankton (132), temperature (8), and zooplankton (22).  A full description is available as an appendix in \citepalias{LehrterIR}.

@@

Local sensitivity analyses showed that \ac{do} was sensitive to perturbations in 105 of 178 parameters (59\% of total) in FishTank (see \cref{fig:sensalltile} for other state variables). Within each parameter category, \ac{do} was sensitive to three parameters for optics (27\% of all optic parameters, \cref{tab:optsens}), eight for organic matter (25\%, \cref{tab:omsens}), 78 for phytoplankton (50\%, \cref{tab:phytosens}), six for temperature (27\%, \cref{tab:tempsens}), and 10 for zooplankton (33\%, \cref{tab:zoopsens}). Although \ac{do} had the greatest sensitivity to parameters in the zooplankton category (as percentage of total), the relative effects varied. Among all parameters, sensitivity values ranged from $L1 = $ $2.29\times 10^{-5}$ for \textit{respg$_{p5}$} (phytoplankton) to $1.24$ for \textit{KG2} (organic matter), whereas average sensitivity among all parameters was $L1 = $ $0.12$.  Within categories, sensitivity ranged from $6.78\times 10^{-4}$ (\textit{astarOMZ}) to $0.43$ (\textit{astar490}) for optics, $0.01$ (\textit{KNO3}) to $1.24$ (\textit{KG2}) for organic matter, $2.29\times 10^{-5}$ (\textit{respg$_{p5}$}) to $0.65$ (\textit{Qc$_{p1}$}) for phytopankton, $9.75\times 10^{-5}$ (\textit{Tref(nospA+nospZ)$_{p6}$}) to $9.97\times 10^{-5}$ (\textit{Tref(nospA+nospZ)$_{p4}$}) for temperature, and $5.75\times 10^{-5}$ (\textit{ZQp$_{z1}$}) to $5.75\times 10^{-5}$ (\textit{ZQp$_{z2}$}) for zooplankton (\cref{fig:sensplo}, bottom).  Average sensitivity values in each category were $L1 = $ $0.17$ for optics, $0.34$ for organic matter, $0.12$ for phytoplankton, $9.89\times 10^{-5}$ for temperature, and $0$ for zooplankton.  Within the six phytoplankton groups, \ac{do} was sensitive to the same parameters within each group although the sensitivity magnitudes varied. Average sensitivity across parameters in each phytoplankton groups showed that \ac{do} was most sensitive to the first and third phytoplankton groups (average $L1 = $ $0.12$, $0.12$), whereas sensitivity to parameters in the remaining groups was much lower (all with average $L1 < 1$). Sensitivity of \ac{do} did not vary considerably between parameters in the two zooplankton groups (average $L1 = $ $0$, $0$ for groups one and two).

Response of \ac{do} to parameter perturbations was not uniform across the time series.  \Cref{fig:sensplo} shows variation for the top parameters within each category.  Because FishTank does not include a spatial component, the estimated \ac{do} trend describes a closed, heterotrophic system where respiration processes eventually remove all \ac{do} from the model space.  The initial decrease in the time series reflects change from the initial conditions outside of the growing seasons (i.e., January, February), the spring/summer increase represents production associated with expected seasonal growth, and the remaining time series from August to the end of year shows complete removal of \ac{do} as respiration processes dominate metabolic activity. Although this time series is not a realistic depiction of an actual system, the biogeochemical model behaves as expected in the absence of the hydrodynamic model.  Accordingly, the interpretation of sensitivity results from the simple model has relevance in an ecological context.  As expected, parameter perturbations had the largest effect during the summer months, although the effects varied.  An inrease in 50\% from the parameter default values generally caused a reduction in \ac{do} during the summer, with the exception of the zooplankton parameter, \textit{ZQp$_{z2}$}, which caused an increase in \ac{do}.  The phytoplankton parameter, \textit{Qc$_{p1}$}, had the largest effect such that the \ac{do} time series was similar to the default output in April/May, whereas a dramatic decrease was observed in the remaining months.  The effects of perturbations early in the time series (January, February) showed similar patterns such that a reduction in \ac{do} was most common, particularly for the optics (\textit{astar490}) and phytoplankton (\textit{Qc$_{p1}$}) parameters.

@@

The identifiability analyses suggested that most parameter subsets exceeded the thresholds of $\gamma = 10, 15$, providing further justification for using selection heuristics for parameter optimization.  Parameter identifiability (as $\gamma$) increased at different rates depending on the parameter category or the number of top parameters that were selected (\cref{fig:identbox}).  By category, identifiability was lowest for parameter subsets in the phytoplankton ($1.7$\% less than $\gamma = 15$, $1.2$\% less than $\gamma = 10$)  and zooplankton categories ($7.9$\%, less than $\gamma = 15$, $4.3$\%, less than $\gamma = 10$), whereas a majority of combinations for temperature were identifiable ($NaN\times 10^{NA}$\% less than $\gamma = 15$, $NaN\times 10^{NA}$\%, less than $\gamma = 10$).  All subset combinations for optics and organic matter parameters had $\gamma < 10$.  All parameter subsets for choosing the top, top two, and top three parameters in each category were identifiable (\cref{fig:identbox}), whereas a majority were identifiable for choosing the top four ($2.7$\% less than $\gamma = 15$, $1.9$\% less than $\gamma = 10$) and top five ($2.7$\% less than $\gamma = 15$, $1.9$\%, less than $\gamma = 10$) parameters.

@@

Each of the three selection heuristics (blocked by parameter category, independent of category, all categories equally) for \ac{do} differed in the number of selected parameters and distribution of parameters within each category (if applicable, \cref{fig:heurist,tab:identall1}).  For the first selection heuristic, all sensitive parameters from the optics ($n = $ $2$, $\gamma =$  $3.5$) and organic matter ($n = $ $3$, $\gamma =$  $10.5$) categories were selected, whereas one were selected for phytoplankton ($\gamma = $ $1$), one for temperature ($\gamma = $ $1$), and one for zooplankton ($\gamma = $ $1$).  For the second selection heuristic, two parameters were selected ($\gamma = $ $8.5$, with a majority from the phytoplankton category, \cref{tab:identall1}).  For the third heuristic, four parameters were selected ($\gamma = $ $12.6$) with equal representation between categories. For the second and third selection heuristics, an individual parameter caused a disproportianate increase in $\gamma$ that forced the selection to stop.  Selection independent of category showed that including \textit{Qc$_{p1}$} caused in increase in $\gamma$ from $8.5$ to $17.5$ and equal selection within cateogies showed that including \textit{ZQc$_{z1}$} caused in increase in $\gamma$ from $12.6$ to $72.2$.  Finally, parameter selection for other state variables (\ac{chla}, \ac{par}, nitrate, ammonium, particulate organic matter, dissolved organic matter, phosphorus) also showed that a limited number of parameters were identifiable (for brevity, only results using the second set of selection heuristics are shown, \cref{fig:heurist_stts,tab:identall2}).  Less than ten parameters were selected for each of the remaining state variables with most selected from the phytoplankton category.  Interestingly, no parameters from the temperature category were selected.

@@

% optics sensitivity
%latex.default(totab, file = "", rowlabel = "Description", caption = cap.val,     caption.loc = "top", rowname = Description, size = tabsize,     label = paste0("tab:", tablab))%
\begin{table}[!tbp]
{\normalsize
\caption{Sensitivities of \ac{do} to perturbation of optics parameters.  Sensitivities are based on a 50\% increase from the default parameter value, where $L1$ and $L2$ summarize differences in model output from the default (see \cref{l1,l2}).  Parameters that did not affect \ac{do} are not shown.\label{tab:optsens}} 
\begin{center}
\begin{tabular}{lllll}
\hline\hline
\multicolumn{1}{l}{Description}&\multicolumn{1}{c}{Parameter}&\multicolumn{1}{c}{Value}&\multicolumn{1}{c}{L1}&\multicolumn{1}{c}{L2}\tabularnewline
\hline
Chla specific absorption at 490 nm&\textit{astar490}&$0.04$&$0.43$&$1.95$\tabularnewline
OMA specific absorption at 490 nm&\textit{astarOMA}&$0.1$&$0.09$&$0.4$\tabularnewline
OMZ specific absorption at 490 nm&\textit{astarOMZ}&$0.1$&$6.78\times 10^{-4}$&$0$\tabularnewline
\hline
\end{tabular}\end{center}}

@@

% organic matter sensitivity
%latex.default(totab, file = "", rowlabel = "Description", caption = cap.val,     caption.loc = "top", rowname = Description, size = tabsize,     label = paste0("tab:", tablab))%
\begin{table}[!tbp]
{\normalsize
\caption{Sensitivities of \ac{do} to perturbation of organic matter parameters.  Sensitivities are based on a 50\% increase from the default parameter value, where $L1$ and $L2$ summarize differences in model output from the default (see \cref{l1,l2}).  Parameters that did not affect \ac{do} are not shown.\label{tab:omsens}} 
\begin{center}
\begin{tabular}{lllll}
\hline\hline
\multicolumn{1}{l}{Description}&\multicolumn{1}{c}{Parameter}&\multicolumn{1}{c}{Value}&\multicolumn{1}{c}{L1}&\multicolumn{1}{c}{L2}\tabularnewline
\hline
turnover rate for OM2A and OM2G&\textit{KG2}&$50$&$1.24$&$6.02$\tabularnewline
turnover rate for OM1A and OM1G&\textit{KG1}&$50$&$0.92$&$4.62$\tabularnewline
half-saturation concentration for O2 utilization&\textit{KO2}&$10$&$0.35$&$1.57$\tabularnewline
O2 concentration that inhibits denitrification&\textit{KstarO2}&$10$&$0.06$&$0.34$\tabularnewline
maximum rate of nitrification per day&\textit{nitmax}&$0.52$&$0.05$&$0.21$\tabularnewline
NH4 rate constant for nitrification&\textit{KNH4}&$1$&$0.04$&$0.17$\tabularnewline
decay rate of CDOM, 1/day&\textit{KGcdom}&$0.01$&$0.03$&$0.12$\tabularnewline
half-saturation concentration for NO3 used in denitrification&\textit{KNO3}&$10$&$0.01$&$0.05$\tabularnewline
\hline
\end{tabular}\end{center}}

@@

% phytoplankton sensitivity
%latex.default(totab, file = "", rowlabel = "Description, Parameter",     caption = cap.val, caption.loc = "top", rgroup = unique(Description),     n.rgroup = as.numeric(table(Description)), rgroupTexCmd = NULL,     rowname = Parameter, size = tabsize, label = paste0("tab:",         tablab))%
\begin{table}[!tbp]
{\normalsize
\caption{Sensitivities of \ac{do} to perturbation of phytoplankton parameters.  Sensitivities are based on a 50\% increase from the default parameter value, where $L1$ and $L2$ summarize differences in model output from the default (see \cref{l1,l2}).  Parameters that did not affect \ac{do} are not shown.  Subscripts show the phytoplankton or zooplankton group that applies for the parameter. Parameters less than the 75\textsuperscript{th} percentile (0.59) for L1 were removed for brevity.\label{tab:phytosens}} 
\begin{center}
\begin{tabular}{llll}
\hline\hline
\multicolumn{1}{l}{Description, Parameter}&\multicolumn{1}{c}{Value}&\multicolumn{1}{c}{L1}&\multicolumn{1}{c}{L2}\tabularnewline
\hline
{phytoplankton carbon/cell}&&&\tabularnewline
~~\textit{Qc$_{p1}$}&$1.35\times 10^{-6}$&$0.65$&$3.16$\tabularnewline
~~\textit{Qc$_{p2}$}&$1.35\times 10^{-6}$&$0.65$&$3.16$\tabularnewline
~~\textit{Qc$_{p3}$}&$1.35\times 10^{-6}$&$0.65$&$3.16$\tabularnewline
~~\textit{Qc$_{p4}$}&$1.35\times 10^{-6}$&$0.65$&$3.16$\tabularnewline
~~\textit{Qc$_{p5}$}&$1.35\times 10^{-6}$&$0.65$&$3.16$\tabularnewline
~~\textit{Qc$_{p6}$}&$1.35\times 10^{-6}$&$0.65$&$3.16$\tabularnewline
\hline
\end{tabular}\end{center}}
